---
title: "Permutations Exercises"
author: "Sean Angiolillo"
date: "3/21/2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We will use the following dataset to demonstrate the use of permutations:

```{r message=FALSE}
library(dplyr)
url <- "https://raw.githubusercontent.com/genomicsclass/dagdata/master/inst/extdata/babies.txt"
filename <- basename(url)
if (!file.exists(filename)) download(url, destfile = filename)
babies <- read.table("babies.txt", header = TRUE)
bwt.nonsmoke <- filter(babies, smoke == 0) %>% select(bwt) %>% unlist 
bwt.smoke <- filter(babies, smoke == 1) %>% select(bwt) %>% unlist
```

## Permutations Exercises #1

We will generate the following random variable based on a sample size of 10 and observe the following difference:

```{r}
N = 10
set.seed(1)
nonsmokers <- sample(bwt.nonsmoke , N)
smokers <- sample(bwt.smoke , N)
obs <- mean(smokers) - mean(nonsmokers)
```

The question is whether this observed difference is statistically significant. We do not want to rely on the assumptions needed for the normal or t-distribution approximations to hold, so instead we will use permutations. We will reshuffle the data and recompute the mean. We can create one permuted sample with the following code:

```{r}
dat <- c(smokers,nonsmokers)
shuffle <- sample( dat )
smokersstar <- shuffle[1:N]
nonsmokersstar <- shuffle[(N + 1):(2 * N)]
mean(smokersstar) - mean(nonsmokersstar)
```

The last value is one observation from the null distribution we will construct. Set the seed at 1, and then repeat the permutation 1,000 times to create a null distribution. What is the permutation derived p-value for our observation?

```{r}
set.seed(1)
null <- replicate(1000, {
  shuffle <- sample( dat )
  smokersstar <- shuffle[1:N]
  nonsmokersstar <- shuffle[(N + 1):(2*N)]
  mean(smokersstar) - mean(nonsmokersstar)
})
( sum( abs(null) >= abs(obs)) + 1 ) / ( length(null) + 1 ) 

```

## Permutations Exercises #2

Repeat the above exercise, but instead of the differences in mean, consider the differences in median `obs <- median(smokers) - median(nonsmokers)`. What is the permutation based p-value?

```{r}
set.seed(1)
obs <- median(smokers) - median(nonsmokers)
null <- replicate(1000, {
  shuffle <- sample( dat )
  smokersstar <- shuffle[1:N]
  nonsmokersstar <- shuffle[(N + 1):(2*N)]
  median(smokersstar) - median(nonsmokersstar)
})
( sum( abs(null) >= abs(obs)) + 1 ) / ( length(null) + 1 ) 
```

