---
title: "Boxplot Exercises"
author: "Sean Angiolillo"
date: "3/21/2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The `InsectSprays` data set measures the counts of insects in agricultural experimental units treated with different insecticides. This dataset is included in R, and you can examine it by typing:

```{r}
head(InsectSprays)
```

## Boxplot Exercises #1

Which spray seems the most effective (has the lowest median)?

```{r message=FALSE}
library(dplyr)
library(ggplot2)
ggplot(InsectSprays, aes(x = spray, y = count)) + geom_boxplot()

InsectSprays %>%
    group_by(spray) %>%
    summarize(n = n(), med_count = median(count)) %>%
    arrange(med_count)
```

* A

* B

* **C**

* D

* E

## Boxplot Exercises #2

Let's consider a random sample of finishers from the New York City Marathon in 2002. This dataset can be found in the UsingR package. Load the library and then load the `nym.2002` dataset.

Use boxplots and histograms to compare the finishing times of males and females. Which of the following best describes the difference?

The boxplot and density plot do a good job of showing how the male and female distributions have a similar shape (slightly right skewed) with the male center shifted to the left (faster times). The histogram reminds us of the much greater number of male finishers.

```{r}
data(nym.2002, package = "UsingR")
ggplot(nym.2002, aes(x = gender, y = time)) + geom_boxplot() + coord_flip()
ggplot(nym.2002, aes(x = time)) + geom_histogram() + facet_wrap(~ gender)
ggplot(nym.2002, aes(x = time, fill = gender)) + geom_density(alpha = 0.5)
```

* Males and females have the same distribution.

* Most males are faster than most women.

* **Male and females have similar right skewed distributions with the former, 20 minutes shifted to the left.**

* Both distribution are normally distributed with a difference in mean of about 30 minutes.
